{\LARGE Basis Choice for Scalar-on-Function Regression \\ with Applications to Near-Infrared Spectroscopy}

### 2022-02-11 Paper/main.tex
One passive voice removed
Changes to interpretation of results and indications

		 Two different error-terms were created by generating an i.i.d. standard normal error term and multiplying it by two error variations $\sigma_e $. The error variations represent different signal-to-noise ratios of the responses to test the methods with low and high amounts of noise in the responses. They are created such that the squared multiple correlation coefficient $R^2 = var(\langle X, f\rangle) / (var(\langle X, f\rangle) + \sigma^2_{e})$ is equal to 0.9 and 0.6. The two error terms are then used to generate two sets of responses for $f \in \{f_1(t), f_2(t)\}$.	
		\begin{equation}
			\begin{split}
				Y_{1,f} & = \langle NIR_{sim}, f\rangle + Z  \biggl\lbrack\frac{var(\langle NIR_{orig}, f\rangle)}{0.9} - var(\langle NIR_{orig}, f\rangle)\biggr\rbrack \\
				Y_{2,f} & = \langle NIR_{sim}, f\rangle + Z  \biggl\lbrack\frac{var(\langle NIR_{orig}, f\rangle)}{0.6} - var(\langle NIR_{orig}, f\rangle)\biggr\rbrack
			\end{split}
		\end{equation}

@@

	\subsubsection{Basis Expansion Regression}
	The following results originate from the \hyperref[basis_exp_transf]{Estimation using Basis-Representation}, in which we transform the smooth curves to perform scalar-on-function regression. Examining the results, it appears that the cross-validated MSPE exhibits a convex behavior in the number of basis functions.

@@

	\paragraph{Monomial Basis}
	For $f_1$, the cross-validated MSPE is decreasing in the number of FPC's and chosen basis functions for both the high and the low signal-to-noise ratio ($4,5,6$ basis functions for $n_{FPC} = 2, 3, 4$). In  $f_2$, we also observe a decreasing MSPE in the number of FPC's, but no clear relationship for the chosen basis functions. The Monomial basis shows the weakest performance for all three bases in each setting for all $n_{FPC}$.
	\vspace{-0.2cm}

	\paragraph{B-spline Basis}
	For $f_1,Y_1$, the MSPE suggests that models with a higher number of FPC's perform better. While the number of basis functions stays at five for $n_{FPC} = 2,3$, it increases to 6 basis functions for $n_{FPC} = 4$. In $f_1,Y_2$, the cross-validated MSPE is only slightly affected by $n_{FPC}$, but lowest for $n_{FPC} = 3$, indicating the possibility of overfitting using $n_{FPC} = 4$. Similar to the two setups with $f_1$, the chosen number of basis functions for $f_2$ is increasing in the number of FPC's (4, 6, 23 for $n_{FPC} = 2, 3, 4$)
	\vspace{-0.2cm}

	\paragraph{Fourier Basis}
	In $f_1,Y_1$, the MSPE decreases in $n_{FPC}$ while the results in $f_1,Y_2$ might indicate overfitting when using $n_{FPC} = 4$. $f_2,Y_1$ displays the greatest relative decline of MSPE in the increasing number of FPC's, but in absolute values, therefore acknowledging the higher noise responses for $f_2,Y2$, the decline of MSPE is similar to the decline observed for $f_2,Y_2$. Both configurations of $f_2$ use 5, 15 and 7 basis functions for $n_{FPC} = 2, 3, 4$.
	\vspace{-0.2cm}

@@

	This convex behavior was not observed for the FPCR where no clear relationship between the basis functions and the number of FPC's could be found. The MSPE of FPCR seems to be comparatively more stable when changing the number of basis functions than for basis expansion regression. It seems like the chosen number of FPC's is more significant. To examine this relationship further, we conducted additional simulations (Appendix \ref{Tables_sim_additional}) for B-spline basis specifications with a large truncation parameter $L \in \{50, 70\}$ for $n_{FPC} \in \{2, \dots, 7 \}$. As in the simulation study, we can find potential signs of overfitting for all setups and both choices of $L$. The additional simulations revealed evidence for a relationship from $n_{FPC} = 3$ onwards. For $n_{FPC} \in \{3, \dots, 7 \}$, the ten-fold cross-validated MSPE is always lower for 50 than for 70 basis functions which might be the first sign of an $L$ that is chosen too large and creates undersmoothing. Following this, it seems that once a sufficient number of basis functions is used to expand the curves, FPCR performs better using a lower number of basis functions. For $n_{FPC} \in \{2, 3, 4 \}$, the settings $f_1,Y_2$ and $f_2,Y_2$ perform worse for these high number of basis functions compared to the main simulation. In contrast, $f_1,Y_1$ and $f_2,Y_1$ show a stronger performance, although the signal-to-noise ratios of the responses are not affecting the construction of the FPC's. This might indicate beneficial attributes of a lower truncation parameter in settings with low signal-to-noise responses.

>>>>

		 We created two different error-terms by generating an i.i.d. standard normal error term and multiplying it by two error variations $\sigma_e $. The error variations represent different signal-to-noise ratios of the responses to test the methods with low and high amounts of noise in the responses. They are created such that the squared multiple correlation coefficient $R^2 = var(\langle X, f\rangle) / (var(\langle X, f\rangle) + \sigma^2_{e})$ is equal to 0.9 and 0.6. The two error terms are then used to generate two sets of responses for $f \in \{f_1(t), f_2(t)\}$.	
		\begin{equation}
			\begin{split}
				Y_{1,f} & = \langle NIR_{sim}, f\rangle + Z  \biggl\lbrack\frac{var(\langle NIR_{orig}, f\rangle)}{0.9} - var(\langle NIR_{orig}, f\rangle)\biggr\rbrack \\
				Y_{2,f} & = \langle NIR_{sim}, f\rangle + Z  \biggl\lbrack\frac{var(\langle NIR_{orig}, f\rangle)}{0.6} - var(\langle NIR_{orig}, f\rangle)\biggr\rbrack
			\end{split}
		\end{equation}

@@

	\subsubsection{Basis Expansion Regression}
	The following results originate from the \hyperref[basis_exp_transf]{Estimation using Basis-Representation}, in which we transform the smooth curves to perform scalar-on-function regression. Examining the results, it appears that the cross-validated MSPE exhibits a convex behavior in the number of basis functions for all bases.

@@

	\paragraph{Monomial Basis}
	For $f_1,Y_1$, the cross-validated MSPE seems to decrease in the number of FPC's and chosen basis functions for both the high and the low signal-to-noise ratio ($4,5,6$ basis functions for $n_{FPC} = 2, 3, 4$). For the noisy responses of $f_1,Y_2$ we find sign of overfitting for $n_{FPC}$ = 4. In  $f_2$, we also observe a decreasing MSPE in the number of FPC's, but no clear relationship for the chosen basis functions. The Monomial basis shows the weakest performance for all three bases in each setting.
	\vspace{-0.2cm}

	\paragraph{B-spline Basis}
	For $f_1,Y_1$, the MSPE suggests that models with a higher number of FPC's perform better. While the number of basis functions stays at five for $n_{FPC} = 2,3$, it increases to 6 basis functions for $n_{FPC} = 4$. In $f_1,Y_2$, the cross-validated MSPE is only slightly affected by $n_{FPC}$, but lowest for $n_{FPC} = 3$, indicating the possibility of overfitting for $n_{FPC} = 4$. Similar to the two setups with $f_1$, the chosen number of basis functions for $f_2$ is increasing in the number of FPC's (4, 6, 23 for $n_{FPC} = 2, 3, 4$)
	\vspace{-0.2cm}

	\paragraph{Fourier Basis}
	In $f_1,Y_1$, the MSPE decreases in $n_{FPC}$ while the results in $f_1,Y_2$ might indicate overfitting when using $n_{FPC} = 4$. Both specifications using $f_1$ select the lowest number of basis  possible. $f_2,Y_1$ displays the greatest relative decline of MSPE in the increasing number of FPC's, but in absolute values, therefore acknowledging the higher noise responses for $f_2,Y2$, the decline of MSPE is similar to the decline observed for $f_2,Y_2$. Both configurations of $f_2$ use 5, 15 and 7 basis functions for $n_{FPC} = 2, 3, 4$. The basis shows the lowest MSPE for all four settings.
	\vspace{-0.2cm}

@@

	This convex behavior was not observed for the FPCR where no clear relationship between the basis functions and the number of FPC's could be found. The MSPE of FPCR seems to be comparatively more stable when changing the number of basis functions than for basis expansion regression. It seems like the chosen number of FPC's is more significant, especially for the two high signal-to-noise ratio settings which display a higher relative decrease of MSPE in $n_{FPC}$. To examine this relationship further, we conducted additional simulations (Appendix \ref{Tables_sim_additional}) for B-spline basis specifications with a large truncation parameter $L \in \{50, 70\}$ for $n_{FPC} \in \{2, \dots, 7 \}$. We can find potential signs of overfitting in all setups for $L = 70$. For $L = 50$, signs of overfitting can be observed for the setups with the noisy responses. The additional simulations revealed evidence for a relationship from $n_{FPC} = 3$ onwards for all setups. For $n_{FPC} \in \{3, \dots, 7 \}$, the ten-fold cross-validated MSPE is always lower for 50 than for 70 basis functions which might be the first sign of an $L$ that is chosen too large and creates undersmoothing. Following this, it seems that once a sufficient number of basis functions is used to expand the curves, FPCR with B-splines performs better using a lower number of basis functions. While a higher number of $n_{FPC}$ decreases the MSPE of $f_1,Y_1$ and $f_2,Y_1$ for $L = 50$ compared to the main simulations, this holds not true for the two setups with noisy responses. This strengthens our hypothesis of the high significance of $n_{FPC}$ for the high signal-to-noise ratio setup. For $n_{FPC} \in \{2, 3, 4 \}$, the additional B-splines simulations perform worse for these high number of basis functions compared to the main simulation.  This might indicate beneficial attributes of a lower truncation parameter in settings with a lower number of $n_{FPC}$.
